---
title: "DAGs propuestas"
author: "Rafael Gutiérrez"
format: 
    html:
     toc: true
     html-math-method: katex
     embed-resources: true
     self-contained-math: true
     df-print: kable
editor: source
editor: visual
---

## DAGs propuestas

```{r}
#| eval: false
install.packages("bnlearn")
```
```{r}
library(bnlearn)
```
```{r}
data=read.csv("",stringsAsFactors = TRUE)
head(data)
```
Para el manejo de la DAG se asignó una clave corta a cada variable:
    A = edad_cat,
    D = dia_viaje,
    EST = estrato,
    V = autos_camionetas,
    L = localidad,
    H = tiempo_cat,     
    O = Entidad,        
    ESC = escolaridad,
    S = sexo,
    T = transporte
```{r}
dag1 = empty.graph(nodes = c("A", "D", "EST", "V", "L", "H", "O", "ESC","S","T"))
```
```{r}
arc_set1 = matrix(c("A","ESC",
                   "ESC","EST",
                   "EST","V",
                   "V","T",
                   "T","H",
                   "S","T",
                   "D","T",
                   "O","L",
                   "L","EST",
                   "L","H"), 
                 byrow = TRUE, ncol = 2,
                 dimnames = list(NULL, c("from","to")))
```
```{r}
arcs(dag1) = arc_set1
```
```{r}
dag1
```
```{r}
graphviz.plot(dag1, shape = "rectangle")
```

```{r}
dag2 = empty.graph(nodes = c("A", "D", "EST", "V", "L", "H", "O", "ESC","S","T"))
```

```{r}
arc_set2 = matrix(c("O","L",
                   "L","EST",
                   "EST","T",
                   "O","H",
                   "L","H",
                   "EST","V",
                   "V","T",
                   "S","T",
                   "A","ESC",
                   "ESC","T",
                   "D","H"), 
                 byrow = TRUE, ncol = 2,
                 dimnames = list(NULL, c("from","to")))
```
```{r}
arcs(dag2) = arc_set2
```
```{r}
dag2
```
```{r}
graphviz.plot(dag2, shape = "rectangle")
```
```{r}
dag3 = empty.graph(nodes = c("A", "D", "EST", "V", "L", "H", "O", "ESC","S","T"))
```
```{r}
arc_set3 = matrix(c("A","ESC",
                   "ESC","EST",
                   "EST","T",
                   "T","H",
                   "S","T",
                   "A","V",
                   "V","T",
                   "O","L",
                   "L","T",
                   "D","T",
                   "L","H"), 
                 byrow = TRUE, ncol = 2,
                 dimnames = list(NULL, c("from","to")))
```
```{r}
arcs(dag3) = arc_set3
```
```{r}
dag3
```
```{r}
graphviz.plot(dag3, shape = "rectangle")
```



## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
